Analysis of Bauer et al. (unpublished) Field experiment: <br> Favourable Conservation Status (FCS)

### 2023-03-01 R/model_check_fcs.Rmd
Update model_check_fcs.Rmd

NUTS sampler is used.

@@

```{r analysis-preparation}
# Chose variables
variables <- c(
      "Intercept",
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
# Subset draws
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior_prior <- m_prior %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior_flat <- m_flat %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
# R hat
rhat1 <- brms::rhat(m_1)
rhat2 <- brms::rhat(m_2)
# NEFF ratio
neff1 <- brms::neff_ratio(m_1)
neff2 <- brms::neff_ratio(m_2)
# Long format of draws
hmc_diagnostics1 <- brms::nuts_params(m_1)
hmc_diagnostics2 <- brms::nuts_params(m_2)
y <- sites$n
# Posterior predictive distribution
yrep1 <- brms::posterior_predict(m_1, draws = 500)
yrep2 <- brms::posterior_predict(m_2, draws = 500)
yrep_prior <- brms::posterior_predict(m_prior, draws = 500)
# Leave-one-out cross validation based on posterior likelihood
loo1 <- brms::loo(m_1, save_psis = TRUE, moment_match = FALSE)
loo2 <- brms::loo(m_2, save_psis = TRUE, moment_match = FALSE)
# Summary statistics
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

@@

```{r prior-predictive-check}
bayesplot::ppc_dens_overlay(yrep_prior[1:500, ])
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$site) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$exposition) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$survey_year_fct) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$target_type) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$seed_density) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$sand_ratio) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$substrate_depth) +
  coord_cartesian(xlim = c(-2, 2))
```

@@

#### Sampling efficency: Rhat (BARG 2.B)

```{r rhat, collapse = TRUE}
bayesplot::mcmc_rhat(rhat1)
bayesplot::mcmc_rhat(rhat1)
```

@@

```{r ess, collapse = TRUE}
bayesplot::mcmc_neff(neff1)
bayesplot::mcmc_neff(neff2)
```

@@

```{r loo}
loo1
loo2
plot(loo1)
plot(loo2)
```

>>>>

Markov Chain Monte Carlo (MCMC) method used with No-U-Turn Sampler
(NUTS) sampler which is an extension of Hamiltonian Monte Carlo (HMC).

@@

```{r analysis-preparation}
# Chose variables
variables <- c(
      "Intercept",
      "b_sand_ratio25",
      "b_sand_ratio50",
      "b_substrate_depth15",
      "b_target_typedry_grassland",
      "b_seed_density8",
      "b_expositionsouth",
      "b_survey_year_fct2019",
      "b_survey_year_fct2020",
      "b_survey_year_fct2021",
      "sd_site__Intercept",
      "sd_site:plot__Intercept",
      "sigma"
    )
# Subset draws
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior_prior <- m_prior %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior_flat <- m_flat %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
# Long format of draws
hmc_diagnostics1 <- m_1 %>% brms::nuts_params()
hmc_diagnostics2 <- m_2 %>% brms::nuts_params()
y <- sites$n
# Posterior predictive distribution
yrep1 <- m_1 %>% brms::posterior_predict(draws = 500)
yrep2 <- m_2 %>% brms::posterior_predict(draws = 500)
yrep_prior <- m_prior %>% brms::posterior_predict(draws = 500)
# Summary statistics
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

@@

```{r prior-predictive-check}
bayesplot::ppc_dens_overlay(y, yrep_prior[1:500, ])
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$site) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$exposition) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$survey_year_fct) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$target_type) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$seed_density) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$sand_ratio) +
  coord_cartesian(xlim = c(-2, 2))
ppc_dens_overlay_grouped(y, yrep_prior[1:500, ], group = sites$substrate_depth) +
  coord_cartesian(xlim = c(-2, 2))
```

@@

#### Sampling efficency: R-hat (BARG 2.B)

```{r rhat, collapse = TRUE}
m_1 %>% brms::rhat() %>% bayesplot::mcmc_rhat()
m_2 %>% brms::rhat() %>% bayesplot::mcmc_rhat()
```

@@

```{r ess, collapse = TRUE}
m_1 %>% brms::neff_ratio() %>% bayesplot::mcmc_neff()
m_2 %>% brms::neff_ratio() %>% bayesplot::mcmc_neff()
```

@@

```{r loo}
loo1 <- m_1 %>% brms::loo(save_psis = TRUE, moment_match = FALSE)
loo2 <- m_2 %>% brms::loo(save_psis = TRUE, moment_match = FALSE)
loo1
loo2
plot(loo1)
plot(loo2)
```
